---
title: "P8105_hw2_xl2851"
output: github_document
---

# problem 1
## 1(1).read and clean the Mr. Trash Wheel sheet
```{r}
library(tidyverse)
library(readxl)
library(dplyr)

trash_data = read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="Mr. Trash Wheel", range = cell_cols("A2:N408")) %>%
  janitor::clean_names() %>%
  select(-x15, -x16,-x17) %>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls, digit=0),    sports_balls=as.integer(sports_balls))
  
```

## 1(2).read and clean precipitation data for 2017 and 2018
```{r}
prec_data_2017=read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2017 Precipitation", skip=1) %>%
  janitor::clean_names() %>%
  mutate(year='2017')%>%
  drop_na()
  
```
```{r}
prec_data_2018=read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2018 Precipitation", skip=1) %>%
  janitor::clean_names() %>%
  mutate(year='2018')%>%
  drop_na()
```

## 1(3).combine datasets and convert month
```{r}
prec_data_join=
  full_join(prec_data_2017, prec_data_2018)%>%
  mutate(month = month.name [month] )
```
(1)There are `r nrow(prec_data_join)` observations in the joint precipitation dataset.

The total precipitation in 2018 is `r sum(pull(prec_data_2018,total))`

Examples of key variables:total, describing the total precipitation in each month in 2017 and 2018.

(2)There are `r nrow (trash_data)` observations in the Mr.Trash Wheel sheet.

The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_data%>% filter(year == 2017), sports_balls))`

Examples of key variables:weight_tons, volume_cubic_yards, plastic_bottles

# problem 2
## 2(1).clean the data in pols_months
```{r}
  pols_data= read_csv("./fivethirtyeight_datasets/pols-month.csv")%>%
janitor::clean_names()%>%
  separate(mon, c("year", "month", "day"))%>% 
    mutate(month=month.abb[as.numeric(month)],
    month=tolower(month),
    president=ifelse(prez_gop, 'gop','dem'))%>% 
    select(-prez_dem, -prez_gop, -day)
```

## 2(2).clean the data in snp.csv
```{r}
snp_data = read_csv(file = "./fivethirtyeight_datasets/snp.csv")%>%
  janitor::clean_names()%>%
  separate(col=date,into= c("month", "day", "year"))%>% 
  mutate(month=month.abb[as.numeric(month)],
           month=tolower(month))%>% 
  select(year, month,close,-day)
```

## 2(3).tidy the unemployment data
```{r}
  unemployment_tidy_data =read_csv(file = "./fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "%_of_unemployment")%>%
  mutate(month=tolower(month),Year=as.character(Year))%>%
  select(year=Year, everything())
    
```

## 2(4).join the 3 datasets
```{r}
pols_snp_join=
  left_join(pols_data, snp_data, by=c("year", "month"))
```

```{r}
unemployment_pols_snp_join=
  left_join(pols_snp_join,unemployment_tidy_data)
```

1. In the cleaned "pols" dataset, there are `r nrow(pols_data)` rows and `r ncol(pols_data)`columns.The year range in this dataset is from `r min(pull(pols_data, year))` to `r max(pull(pols_data, year))`.

key variables:

(1)president:indicator of whether the president was republican or democraticon the associated date (1 = yes, 0 = no)

(2)gov_gop,sen_gop,rep_gop:the number of republican governors,senators,representatives on the associated date

(3)gov_dem,sen_dem,rep_dem:the number of democratic governors,senators,representatives on the associated date

2. In the cleaned "snp" dataset, there are `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.The year range in this dataset is from `r min(pull(snp_data, year))` to `r max(pull(snp_data, year))`.

key variables:

(1)close: the closing values of the S&P stock index on the associated date

3. In the cleaned "unemployment" dataset, there are `r nrow(unemployment_tidy_data)` rows and `r ncol(unemployment_tidy_data)` columns.The year range in this dataset is from `r min(pull(unemployment_tidy_data, year))` to `r max(pull(unemployment_tidy_data, year))`.

key variables:

(1)"% of unemployment": the created variable "percentage of unemployment" is a key factor to reflect the administration of a president.

4. In the cleaned joint dataset, there are `r nrow(unemployment_pols_snp_join)` rows and `r ncol(unemployment_pols_snp_join)` columns.The year range in this dataset is from `r min(pull(unemployment_pols_snp_join, year))` to `r max(pull(unemployment_pols_snp_join, year))`.

key variables:

(1)president:indicator of whether the president was republican or democraticon the associated date (1 = yes, 0 = no)

(2)gov_gop,sen_gop,rep_gop:the number of republican governors,senators,representatives on the associated date

(3)gov_dem,sen_dem,rep_dem:the number of democratic governors,senators,representatives on the associated date

(4)close: the closing values of the S&P stock index on the associated date

(5)"% of unemployment": the created variable "percentage of unemployment" is a key factor to reflect the administration of a president.

# problem 3
## 3(1).load and tidy the dataset
```{r}
popbabynames_data = read_csv(file = "./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    childs_first_name=str_to_lower(childs_first_name),
    gender=str_to_lower(gender),
     ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP"= "BLACK NON HISPANIC","WHITE NON HISP" = "WHITE NON HISPANIC")
   )%>%
   distinct()
```

## 3(2).rank in popularity of the name "olivia"
```{r}
  filter(popbabynames_data, gender == "female", childs_first_name == "olivia")%>%
  select(ethnicity, rank, year_of_birth)%>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable()
```

## 3(3).Produce a similar table showing the most popular name among male children over time
```{r}
    filter(popbabynames_data, gender == "male", rank =="1")%>%
  select(gender,ethnicity, rank, childs_first_name,year_of_birth)%>%
pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% knitr::kable()
```

## 3(4).scatter plot
```{r}
scatterplot_df=filter(popbabynames_data, gender == "male", year_of_birth=="2016", ethnicity=="WHITE NON HISPANIC")

ggplot(scatterplot_df, aes(x = rank, y = count)) + geom_point()+
  labs(
    title = "The first name of male, white non-hispanic children born in 2016",
    x = "The rank in popularity of that name (rank)",
    y = "The number of children with a name (count)")
```

