---
title: "P8105_hw2_xl2851"
output: github_document
---

# problem 1
## read and clean the Mr. Trash Wheel sheet
```{r}
library(tidyverse)
library(readxl)
library(dplyr)

trash_data = read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="Mr. Trash Wheel", range = cell_cols("A2:N408")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls, digit=0), sports_balls_new=as.integer(sports_balls))
  
```

## read and clean precipitation data for 2017 and 2018
```{r}
prec_data_2017=read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2017 Precipitation", skip=1) %>%
  janitor::clean_names() %>%
  mutate(year='2017')%>%
  drop_na()
  
```
```{r}
prec_data_2018=read_excel(
  "./Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet="2018 Precipitation", skip=1) %>%
  janitor::clean_names() %>%
  mutate(year='2018')%>%
  drop_na()
```

## combine datasets and convert month
```{r}
prec_data_join=
  full_join(prec_data_2017, prec_data_2018)%>%
  mutate(month = month.name [month] )
```
There are `r nrow(prec_data_join)` observations in the joint precipitation dataset.

There are `r nrow (trash_data)` observations in the Mr.Trash Wheel sheet.

Examples of key variables:

The total precipitation in 2018 is `r sum(pull(prec_data_2018,total))`

The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_data,sports_balls))`


# problem 2
## first,clean the data in pols_months
```{r}
pols_data= read_csv("./fivethirtyeight_datasets/pols-month.csv")%>%
  separate(mon, c("year", "month", "day"))%>% 
    mutate(month=month.name[month_numeric=as.numeric(month)],president='gop','dem')%>% select(-prez_dem, -prez_gop, -day)
```

## second, clean the data in snp.csv
```{r}
snp_data = read_csv(file = "./fivethirtyeight_datasets/snp.csv")%>% 
 separate(col=date,into= c("month", "day", "year"))%>% 
    mutate(month=month.name[month_numeric=as.numeric(month)])%>% 
  select(-day)%>%
  select("year", "month","close")
```

## third, tidy the unemployment data
```{r}
unemployment_data = read_csv(file = "./fivethirtyeight_datasets/unemployment.csv")

```




